Runtime Feedback in a Meta-Tracing JIT for Efficient Dynamic Languages

### 2011-04-13 talk/icooolps2011/paper.tex
footnotize a paragraph

Instances were optimized making the assumption that the total number of
different instance layouts is small compared to the number of instances. For classes we
will make an even stronger assumption. We simply assume that it is rare for
classes to change at all. This is not totally reasonable (sometimes classes contain
counters or similar things) but for this simple example it is good enough.

@@

The first problem that needs to be solved is that Python supports (multiple)
inheritance. Therefore looking up a method in a class needs to consider all the
classes in the
whole method resolution order. This makes the versioning of classes more
complex. If a class is changed its version changes. At the same time, the
versions of all the classes inheriting from it need to be changed as well,
recursively. This makes class changes expensive, but they should be rare.  On the
other hand, a method lookup in a complex class hierarchy is as optimized in the
trace as in our object model here.

A downside of the versioning of classes that we haven't yet fixed in PyPy, is
that some classes \emph{do} change a lot. An example would be a class that keeps a
counter of how many instances have been created so far. This is very slow right
now, but we have ideas about how to fix it in the future.

Another optimization is that in practice the shape of an instance is correlated
with its class. In our code above, we allow both to vary independently.
In PyPy's Python interpreter we act somewhat more cleverly. The class of
an instance is not stored on the instance itself, but on the map. This means
that we get one fewer promotion (and thus one fewer guard) in the trace,
because the class doesn't need to
be promoted after the map has been.

>>>>

Instances were optimized making the assumption that the total number of
different instance layouts is small compared to the number of instances. For classes we
will make an even stronger assumption. We simply assume that it is rare for
classes to change at all. This is not totally reasonable (sometimes classes contain
counters or similar things) but for this simple example it is good
enough.\footnote{There is a more complex variant of class versions that can
accommodate class fields that change a lot better.}

@@

The first problem that needs to be solved is that Python supports (multiple)
inheritance. Therefore looking up a method in a class needs to consider all the
classes in the
whole method resolution order. This makes the versioning of classes more
complex. If a class is changed its version changes. At the same time, the
versions of all the classes inheriting from it need to be changed as well,
recursively. This makes class changes expensive, but they should be rare.  On the
other hand, a method lookup in a complex class hierarchy is as optimized in the
trace as in our simple object model above.

Another optimization is that in practice the shape of an instance is correlated
with its class. In our code above, we allow both to vary independently.
Therefore we store the class of an instance on the map in PyPy's Python
interpreter. This means that we get one fewer promotion (and thus one fewer
guard) in the trace, because the class doesn't need to be promoted after the
map has been.
